Optimizing DTW-Based Audio-to-MIDI Alignment and Matching

### 2015-09-21 paper.tex
Add accuracy experiment section
Also switched to \ninept in the vague dream that this paper will fit.

\begin{document}

\maketitle

@@

Given the framework described above for generating a synthetic alignment dataset,

Short overview of Bayesian optimization

Parameter space (including multiplicative penalty)

Random trials

Discussion of best-performing aligners; also best aligner with beats

>>>>

\begin{document}
\ninept
\maketitle

@@

Given a dataset of MIDI/audio pairs where we know a priori the correct alignment, we can rapidly evaluate the performance of a given alignment scheme by aligning each pair and computing the mean absolute alignment error.
In the present setting, computing the mean error quantifies the extent to which the alignment was able to remove the timing distortions (random warping and cropping) described in Section \ref{sec:synthetic}.
When the alignment has failed for a portion of the song, the error between the mapped times and the correct times may be very large, so we clip the mean error to $.5$ seconds (which essentially denotes an error threshold above which all local alignment discrepencies are equally incorrect).
This results in the following metric for computing the accuracy of a given alignment scheme on a given MIDI/audio pair:
$$
\frac{1}{L}\sum_{i = 1}^{L} \min(|t_X[p[i]] - \hat{t}_X[q[i]]|, .5)
$$
where $\hat{t}_X$ is the ground-truth correct aligned timing.
To compute the accuracy of a given alignment system, we can compute the mean of this metric accross all pairs in the dataset.

The ability to rapidly evaluate an alignment system allows us to perform large-scale comparisons of different parameter settings.
To decide which parameter settings to evaluate, we use Bayesian optimization, which treats the process of tuning parameters to achieve different accuracies as a Gaussian process.
Using this formulation, Bayesian optimization can automatically propose new alignment systems based on the performance of previously evaluated systems.
As a performance objective, we used the mean of mean absolute errors across the entire ``easy'' alignment dataset.
A more thorough discussion can be found in \cite{snoek2012practical}, which is the framework we used in this experiment.

Before applying Bayesian optimization, we must define a parameter space to search over.
Based on the design choices outlined in Section \ref{sec:dtw}, we chose the following parameter space:
\begin{description}[topsep=1pt,itemsep=-1pt,leftmargin=5pt]
\item[Feature representation:] We used either chroma vectors or constant-Q spectra.
The constant-Q spectra spanned 4 octaves, starting from MIDI note C3 (65.4 Hz) with 12 bins per octave.
In preliminary experiments, we found that all of the best-performing alignment systems used log-magnitude features regardless of whether chroma vectors or constant-Q spectra were used, so we computed log-magnitude features in all experiments.
\item[Time scale:] We either computed feature vectors every 46 milliseconds or utilized beat-synchronous features.
\item[Normalization:] We optionally z-scored the feature vectors, and normalized feature vectors by their $L^1$, $L^2$, $L^\infty$ (max) norm, or not at all.
\item[Penalty:] For the penalty, we optimized a scale in $[0, 3]$ to apply to the median distance between all pairs of feature vectors in $X$ and $Y$.
Using the median distance made this penalty adaptive to different feature representations and normalization schemes.
\item[Gully and band path constraint:] We allowed the gully to take any value in $[0, 1]$ and optionally enforced the band path constraint (Sakoe-Chiba band).
\end{description}
This space subsumes most of the systems discussed in Section \ref{sec:dtw}.
All feature extraction was realized with \texttt{librosa} \cite{mcfee2015librosa}.

When performing Bayesian optimization, it is helpful to ``seed'' the optimization with objective values for many randomly-chosen parameter settings to ensure that the optimization thoroughly explores the possible parameter space.
We computed the accuracy for 10,000 randomly configured alignment systems and seeded the optimizer with the 100 systems which achieved the lowest mean error.
In order to avoid local minima in the parameter space, we carried out 10 separate Bayesian optimization runs with 100 trials each, resulting in 1,000 total trials.

The best-performing alignment system achieved an objective value of $0.0181$, meaning that the average absolute error across the entire dataset was about 18 milliseconds.
This is both within the upper range of the auditory temporal resolution and less than half of the time-scale resolution used for non-beat-synchronous feature vectors, so attaining a higher accuracy is likely unrealistic.
As a high-level overview of the results, there were 34 systems which achieved an objective value less than $0.0185$; none of these systems used beat-synchronization or a path constraint and all of them used log-magnitude constant-Q spectra as a feature representation.
All of the best-performing systems set the penalty median scale close to 1 and the ``gully'' parameter was also uniformly close to 1.
Almost all of these systems used $L^2$ normalization (resulting in a cosine distance for local feature comparisons); a few used $L^1$ normalization.
Whether z-scoring was applied or not varied uniformly across these systems.
